Are populations like a circuit? The relationship between isolation by distance and isolation by resistance

### 2018-10-05 isolation_by_coalescence.tex
begun putting in other sims

\paragraph{Continuous Landscape with Biased Migration}
The third situation that we will discuss is where migration is biased.
We will be using a square landscape as in the first case, 
but the offspring locations of offspring are not centered around the parent.
Instead, the mean of the distribution of the locations
is one tenth of a standand deviation (or 0.05\% of the edge length of the landscape) 
up and to the right of their parent's location.
This results in reverse time gene flow down and to the left 
as in the biased asymmetric graph in Section \ref{sec:biased_migration}.
In terms of population genetics, 
this is very similar to the case where a population is expanding into new territory 
since in both cases, 
offspring tend to be further from the edge of the expansion than their parents.
The difference is that, in our case, 
the population is constrained to be a bounded distance 
away from the edge of the ``expansion''.
In both situations, individuals at the edge may have disproportionately more offspring,
leading to low genetic diversity and strong drift.
This results in substantially more noise in the isolation by distance plots 
as seen in Figure \ref{fig:ibd_comp}. 
The horizontal lines in that figure are likely from comparing genomes of individuals from
two families that have recently increased in size and geographical spread, 
resulting in many pairs of genomes with similar genetic distance 
but varying geographic distance.

@@

\paragraph{The effect of history}
The models we discuss here assume that population sizes and migration rates
have been constant on the time scale given by the within-species coalescent time.
This is rarely true in practice \citep{avise,barton}.
However, geographic differences in mean relatedness
are established on a shorter time scale --
the time scale over which a lineage, moving randomly across the landscape,
``forgets'' where it started
(i.e., the mixing time of the Markov chain).
This is the time it takes for the uncertainty in lineage movement
to reach the scale of the species range.
For instance,
the width of the tortoise range is roughly 500km,
and if mean dispersal distance is 10km (a high estimate),
the standard deviation of ancestor location $t$ generations ago is $10\sqrt{t}$,
so the mixing time of tortoise lineages across the modern landscape,
not accounting for barriers, is of order 2500 generations,
or about 50,000 years.
The landscape of the Mojave has changed substantially over tihs time \citep{pleistocene_mojave},
which suggests that models incorporating change over time 
may be required to model modern tortoise diversity.
However, the effects of recent history are strongest,
so landscapes estimated assuming constant populations
may give a reasonable picture of the landscape averged over recent times.

>>>>

\paragraph{Continuous Landscape with Biased Migration}
We simulated several situations with an effective bias in migration on a continuous landscape.
First, the ``biased'' situation, was a square landscape as above,
but with mean offspring location one tenth of a standard deviation up and to the right of the parent's location.
This produces reverse-time gene flow down and to the left.
Second, we simulated a ``valley'' in a $3 \times 5$ landscape as above, 
in which fecundity of individuals in the middle 1/4 of the landscape was reduced by 50\%.
This produces gene flow ``uphill'', away from the ridge,
and thus an effective barrier across the ridge.
Third, we simulated a recent ``expansion'',
in which the middle 4/5 of the landscape was uninhabitable 400,000 generations,
until 500 generations before the present day.
The resulting expansion produces gene flow back away from the center,
similar to the ``valley''.
\textbf{MAKE FIGURE OF THESE?}
\textbf{Merge with berow.}

The third situation that we will discuss is where migration is biased.
We will be using a square landscape as in the first case, 
but the offspring locations of offspring are not centered around the parent.
Instead, the mean of the distribution of the locations
is one tenth of a standand deviation (or 0.05\% of the edge length of the landscape) 
up and to the right of their parent's location.
This results in reverse time gene flow down and to the left 
as in the biased asymmetric graph in Section \ref{sec:biased_migration}.
In terms of population genetics, 
this is very similar to the case where a population is expanding into new territory 
since in both cases, 
offspring tend to be further from the edge of the expansion than their parents.
The difference is that, in our case, 
the population is constrained to be a bounded distance 
away from the edge of the ``expansion''.
In both situations, individuals at the edge may have disproportionately more offspring,
leading to low genetic diversity and strong drift.
This results in substantially more noise in the isolation by distance plots 
as seen in Figure \ref{fig:ibd_comp}. 
The horizontal lines in that figure are likely from comparing genomes of individuals from
two families that have recently increased in size and geographical spread, 
resulting in many pairs of genomes with similar genetic distance 
but varying geographic distance.

@@

\paragraph{The effect of history}
The models we discuss here assume that population sizes and migration rates
have been constant on the time scale given by the within-species coalescent time.
This is rarely true in practice \citep{avise,barton}.
However, geographic differences in mean relatedness
are established on a shorter time scale --
the time scale over which a lineage, moving randomly across the landscape,
``forgets'' where it started
(i.e., the mixing time of the Markov chain).
This is the time it takes for the uncertainty in lineage movement
to reach the scale of the species range.
For instance,
if the width of a species' range is roughly 500km,
and if mean dispersal distance is 10km,
the standard deviation of ancestor location $t$ generations ago is $10\sqrt{t}$,
so the mixing time of a lineages across the landscape,
not accounting for barriers, is of order 2500 generations,
or about 50,000 years.
The landscape of many areas has changed substantially over this time, % \citep{pleistocene_mojave},
which suggests that models incorporating change over time 
may be required to model modern tortoise diversity.
However, the effects of recent history are strongest,
so landscapes estimated assuming constant populations
may give a reasonable picture of the landscape averged over recent times.
